---
title: "Text Prediction Algorithm"
author: "Matiki"
date: "January 24, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Coursera Data Science Capstone Project

The final project of the Data Science Specialization on Coursera.org involves 
creating a shiny web application that takes a small amount of text as input,
and predicts the next word as output. The function is similar to many text 
messaging apps found on modern smart phones.

Text mining, natural language processing, data wrangling, and all analyses and 
model building were completed using common R packages.

## How it works:

The app was written on a home computer running Windows with an Intel i3 processor 
and 2 Gb of RAM, so a small random sampling of the data was used in order to be
able to run properly. 

The data sample was 'tokenized' into N-grams, or sequences of N words found in a
given text corpus, and aggregated into frequency data frames, through which the 
algorithm searches for terms like a dictionary. N-grams up to 6 were used.

The algorithm works by taking up to the last five words given as input, and 
searching for a match in the 6-gram data frame. If it finds a match, it prints 
out the 6th word. If no match is found, the app looks at only the last four words
given, and searches in the 5-gram data frame. The app continues like this until a 
match is found.

## Example

```{r }

```

## Links:

```{r }
```

